Finite temperature thresholds for stabilizer codes with few measurements

### 2017-04-13 IsingECC/manuscript_ising_ecc.tex
Add files via upload
updated versions

\subsection{Fusion Probability Calculation}

@@

\begin{align}
\label{eq:bayes}
&P(d_1:d_2 | d_1^{x_1,t_1} \wedge d_2^{x_2,t_2}) = \\ \notag
&\frac{P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) P(d_1 : d_2)}{P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2}}
\end{align}

The individual terms in the right hand side of equation \ref{eq:bayes} are straightforward to interpret.  $P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2)$ represents the probability that two measured defects would be at spacetime coordinates $(x_1,t_1)$ and $(x_2,t_2)$ given that they are indeed a pair.  This can be related to the probability that a one dimensional diffusion process with diffusion constant $D$ will have performed an excursion of length $|x_2-x_1|$ in time $|t_2-t_1|$, and is equal to $1 - \rm{erf}(|x_2-x_1|/(2\sqrt{D |t_2 - t_1|}))$ for a random walk with diffusion constant $D$.  This represents the probability that a random walk will not perform an excursion of distance $|x_2-x_1|$ in a time $t_2 - t_1$.  For our analysis, we will choose $D\propto \gamma_0$.  The exact correspondence between $D$ and $\gamma_0$ depends on the details of the error correction algorithm itself, so, in practice, we treat the constant of proportionality as an empirically tuned parameter.

$P(d_1 : d_2)$ represents the probability that two measured defects, $d_1$ and $d_2$, are in fact a pair.  We decompose this into two pieces: a combinatorial piece, and a dynamical piece.

Finally, $P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})$ is the probability that two defects are measured, one at $(x_1,t_1)$, and the other at $(x_2,t_2)$.  This can be decomposed:

The first term we calculated above.  The second term represents the probability of two measurement events, conditioned on those events \emph{not} being part of a pair.  This is essentially the probability that two independent measurement events have occurred, which is approximately the probability that two independent creation events have occurred (assuming defects are measured suitably efficiently).  For a suitably large system at moderately low temperature, this probability can be estimated as $\propto \frac{|t_2-t_1|}{(L \gamma_+)^{-2}}:=\delta(L,T,\Delta t)$.

@@

 Suppose that (a) defects are efficiently trapped by the measurement patches, and (b) temperature is low so that there is $\rho<<1$ defect pair per measurement patch.  In this regime, condition (a) guarantees that defects don't escape measurement patches.  Then, the only error pathway is a sequence of incorrect, but nontheless ``plausible'' error correction operations, whereby defects are paired with defects that are not their own pair.

@@

 To estimate the rate at which these spurious error-correction events occur, we first consider a slightly altered model.  Suppose that a single pair of defects is in a system, and that no new pairs will be introduced.  Further, suppose that there's an error process that occasionally increases the distance between this pair by an amount $2\sqrt{\gamma_0 \delta t}$, for hopping rate $\gamma_0$ and for system time $\delta t$.  Note that this means the error process acts on longer distance scales as system time increases.  The particular form of this error process is chosen to mimic a spurious error correction event, where a pair of defects is ``corrected'', but this pair of defects was not a real pair.  As time increases, the characteristic distance over which this error process could occur also increases, in accordance with the typical pair-wise separation between two defects performing a random walk.  This typical distance is exactly the factor used by the error correction algorithm to determine if a pair of defects should be corrected or not.

@@

@@

\begin{figure}
\begin{center}
\scalebox{1}{\includegraphics[width=1.0\columnwidth]{LifetimeVsTemperatureVsEnergy.pdf}}
\end{center}
\caption{Here we plot the lifetime of the system as a function of temperature for a variety of energy scales $\Delta$ (see Eq. \ref{eq:isingam}).}
\label{fig:LifetimeVsTemperatureVsEnergy}
\end{figure}

Finally, in Fig. \ref{fig:LifetimeVsTemperatureVsEnergy}, we plot the lifetime of the system in the presence of the protocol at various energy scales $\Delta$.  By contrasting Figures \ref{fig:LifetimeVsMeasurementFraction} and \ref{fig:LifetimeVsTemperatureVsEnergy}, one can deduce the relative benefits of error suppression via more measurement resources versus error suppression via reservoir engineering (i.e., a larger gap to excitation).

@@

\begin{figure}
\begin{center}
\scalebox{1}{\includegraphics[width=1.0\columnwidth]{tc_fig.pdf}}
\end{center}
\caption{Here we sketch one possible $m=3/4$ geometry for the measurement rails for a realization of our protocol.  More sparse geometries can be realized simply by moving the rails of measurement farther apart.  Measured sites are in light blue, and vertex locations for the toric code are circles.  Spin variables (not pictured) reside directly between neighboring vertices.}
\label{fig:tc_fig}
\end{figure}

@@

We have provided numerical and theoretical evidence of a few-measurement error correction protocol for a stabilizer code with string-like excitations.  The primary technical innovation of our algorithm is the bayesian decoding scheme for pairing defects, sketched in Sec. \ref{sec:ecc_alg}.

>>>>

\subsection{Fusion Probability Calculation}
\label{sec:bayes_prob_calc}

%************************TRIM SUBSTANTIALLY AND PUT IN APPENDIX

@@

\begin{align}
\label{eq:bayes}
&P(d_1:d_2 | d_1^{x_1,t_1} \wedge d_2^{x_2,t_2}) = \\ \notag
&\frac{P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) P(d_1 : d_2)}{P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})}
\end{align}

The individual terms in the right hand side of equation \ref{eq:bayes} are straightforward to interpret.  $P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2)$ represents the probability that two measured defects would be at spacetime coordinates $(x_1,t_1)$ and $(x_2,t_2)$ given that they are indeed a pair.  This can be related to the probability that a one dimensional diffusion process with diffusion constant $D$ will have performed an excursion of length $|x_2-x_1|$ in time $|t_2-t_1|$, and takes the form,

\begin{align}
\label{eq:bayesformula}
P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) = 1 - \rm{erf}\bigg(\frac{|x_2-x_1|}{2\sqrt{D |t_2 - t_1|}}\bigg) 
\end{align}

for a random walk with diffusion constant $D$.  This represents the probability that a random walk will not perform an excursion of distance $|x_2-x_1|$ in a time $t_2 - t_1$.  For our analysis, we will choose $D\propto \gamma_0$.  The exact correspondence between $D$ and $\gamma_0$ depends on the details of the error correction algorithm itself, so, in practice, we treat the constant of proportionality as an empirically tuned parameter.

$P(d_1 : d_2)$ represents the probability that two measured defects, $d_1$ and $d_2$, are in fact a pair.  Finally, $P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})$ is the probability that two defects are measured, one at $(x_1,t_1)$, and the other at $(x_2,t_2)$.  As we discuss in Appendix \ref{sec:bayesdecoder}, these factors are not as important for the decoding scheme as Eq. \ref{eq:bayesformula}.  In practice, we find that using the expression from Eq. \ref{eq:bayesformula} alone is sufficient to provide resilient error correction.  We defer further discussion to the appendix.

@@

 Suppose that (a) defects are efficiently trapped by the measurement patches, and (b) temperature is low so that there is $\rho<<1$ defect pair per measurement patch.  In this regime, condition (a) guarantees that defects don't escape measurement patches.  Then, the only error pathway is a sequence of incorrect, but nonetheless ``plausible'' error correction operations, whereby defects are paired with defects that are not their own pair.

%************************Make explicit here that it's a heuristic argument.  Because the error processes in our model are difficult to etc. etc. etc.

@@

 Treating the full dynamics of errors is difficult, so we first consider a different, simpler error dynamics.  First, we argue how, in the presence of this simpler error model, our algorithm provides an exponentially long lifetime in system size.  Then, we argue how this simpler error model is actually more pessimistic with respect to possible error processes than the true error dynamics---that is, we expect the real model to have a \emph{better} lifetime scaling than would be expected from the simple error model.

 The simple error model is defined as follows:  suppose that a single pair of defects is in a system, and that no new pairs will be introduced.  Further, suppose that there's an error process that occasionally increases the distance between this pair by an amount $2\sqrt{\gamma_0 \delta t}$, for hopping rate $\gamma_0$ and for system time $\delta t$.  Note that this means the error process acts on longer distance scales as system time increases.  The particular form of this error process is chosen to mimic a spurious error correction event, where a pair of defects is ``corrected'', but this pair of defects was not a real pair.  As time increases, the characteristic distance over which this error process could occur also increases, in accordance with the typical pair-wise separation between two defects performing a random walk.  This typical distance is exactly the factor used by the error correction algorithm to determine if a pair of defects should be corrected or not.

@@

%************************PUT IN TERMS OF SYSTEM SIZE
%************************MAYBE REMOVE INTERMEDIATE SYSTEM SIZES - remove 12,16,24 from Fig. 2 or maybe just 4,8,16,32.
%************************Mush together figure 3 and 4 maybe?
%************************Lines in fig. 5 and fig. 7 to guide the eye
%************************include m and system sizes in figure captions
%************************maybe have a plot for Delta critical temperature scaling
%************************make the toric code figure sparser perhaps?

@@

\begin{figure}
\begin{center}
\scalebox{1}{\includegraphics[width=1.0\columnwidth]{LifetimeVsTemperatureVsEnergy.pdf}}
\end{center}
\caption{Here we plot the lifetime of the system as a function of temperature for a variety of energy scales $\Delta$ (see Eq. \ref{eq:isingham}).}
\label{fig:LifetimeVsTemperatureVsEnergy}
\end{figure}

%************************EXPAND THIS PARAGRAPH A LITTLE

Finally, in Fig. \ref{fig:LifetimeVsTemperatureVsEnergy}, we plot the lifetime of the system in the presence of the protocol at various energy scales $\Delta$.  By contrasting Figures \ref{fig:LifetimeVsMeasurementFraction} and \ref{fig:LifetimeVsTemperatureVsEnergy}, one can deduce the relative benefits of error suppression via more measurement resources versus error suppression via hamiltonian engineering (i.e., a larger gap to excitation).

@@

\begin{figure}
\begin{center}
\scalebox{1}{\includegraphics[width=1.0\columnwidth]{tc_fig_2.pdf}}
\end{center}
\caption{Here we sketch one possible $m=63/144$ (for a 12x12 unit cell) geometry for the measurement rails for a realization of our protocol.  More sparse geometries can be realized simply by moving the rails of measurement farther apart.  Measured sites are in light blue, and vertex locations for the toric code are circles.  Spin variables (not pictured) reside directly between neighboring vertices.}
\label{fig:tc_fig}
\end{figure}

@@

We have provided numerical and theoretical evidence of a few-measurement error correction protocol for a stabilizer code with string-like excitations.  The primary technical innovation of our algorithm is the bayesian decoding scheme for pairing defects, sketched in Sec. \ref{sec:ecc_alg}.  So long as an appropriate geometry of measurement devices is in place, and so long as defects undergo diffusive motion via coupling to a thermal bath, this scheme can be extended to higher dimensional stabilizer systems like the toric code, as demonstrated in Sec. \ref{sec:tc_alg}.

This scheme complements what is known about decoders in the presence of \emph{noisy} measurements.  Figures \ref{fig:LifetimeVsMeasurementFraction} and \ref{fig:CriticalTempVsMeasurementFraction} demonstrate how a reduction in the measurement fraction in the lattice corresponds to a concomittant decrease in the threshold temperature, similar to how thresholds are known to be reduced when increasing the noise on measurements.

More fundamentally, our algorithm can be understood as an entropy reduction scheme.  Configurations which give rise to errors in the encoded subspace are exponentially suppressed as system size is made larger.  This is in contrast with ``energetic'' suppression---that is, suppression via cooling or by widening the gap to excitations $\Delta$.  This tradeoff between entropic and energetic contributions is depicted in Figures \ref{fig:LifetimeVsTemperatureVsEnergy} and \ref{fig:LifetimeVsMeasurementFraction}.  Depending on the resource requirements of a particular architecture, the threshold temperature can be tuned either by engineering a larger gap, $\Delta$, or by changing the number of measurements used.  In practice, this will depend on the lowest effective temperature available, the maximum measurement rate, $\chi$, as well as the practical difficulty of employing more measuring devices.

%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This material is based upon work supported by DARPA under Grant No. 3854-UCB-AFOSR-0041 and by the National Science Foundation under Grants No. PIF-0803429 and No. CHE-1213141.  CDF was supported by the NSF Graduate Research Fellowship under Grant DGE-1106400 and by the DOE Office of Science Graduate Student Research (SCGSR) program under contract number DE‐SC0014664.
%%%%%%%%%%%%%%%%%%%%%%

\appendix
%%%%%%%%%%%%%%%%%%%%%%
\section{Bayesian Decoding for the Ising Model}
\label{sec:bayesdecoder}
%%%%%%%%%%%%%%%%%%%%%%

In this section, we provide further discussion of Eq. \ref{eq:bayes}, as well as analytic and numerical arguments for how it can be more simply approximated.

First, we decompose $P(d_1 : d_2)$ into two pieces: a combinatorial piece, and a dynamical piece.

As mentioned in Sec. \label{sec:bayes_prob_calc}, $P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})$ is the probability that two defects are measured, one at $(x_1,t_1)$, and the other at $(x_2,t_2)$.  This can be decomposed:

The first term is precisely Eq. \ref{eq:bayesformula}.  The second term represents the probability of two measurement events, conditioned on those events \emph{not} being part of a pair.  This is essentially the probability that two independent measurement events have occurred, which is approximately the probability that two independent creation events have occurred (assuming defects are measured suitably efficiently).  For a suitably large system at moderately low temperature, this probability can be estimated as $\propto \frac{|t_2-t_1|}{(L \gamma_+)^{-2}}:=\delta(L,T,\Delta t)$.
